Kids Basic Mutual Exclusivity

### 2019-03-29 stats/kids_basic_me.Rmd
kids novelty exp

```{r import data, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)

## importing data
files <- dir("~/Work/Spin/raw_data/kids/basic_me/")

@@

# length(unique(raw_data$subid))
# 
d_raw <-raw_data %>%
  mutate(test_date = as.Date(test_date))
# 
log <- read_excel("../../../SPIN-subject_log.xlsx", 1)%>%
  filter(Condition == "me")%>%
  select(subid,experimenter,keep_drop,sex,dob)
# 
# length(unique(log$subid))
# 
# 
# sum(log$keep_drop=="drop")
# 
d <- left_join(d_raw,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(dob = as.numeric(dob),
         dob = as.Date(dob,origin = "1899-12-30"),
         age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))
# 
# length(unique(d$subid))

#write_csv(d, "data/kids_inf_2_data.csv")

@@

```{r filler by age, echo= FALSE, warning= FALSE}
pf1 <- d %>%
      filter(trial == "filler1" | trial == "filler2")%>%
  group_by(control,subage, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = pf1, aes(x = control, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = pf2, aes(x = control, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Plot test trials by age bin

```{r plot by age, echo= FALSE}

p1 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(control,subage, subid) %>%
  summarise(correct = mean(correct))

p2 <- p1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p1, aes(x = control, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = control, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

@@

```{r plot by numerical age, echo = F}
p5 <- d %>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  group_by(age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p5, aes(x = age_num, y = correct)) +
  geom_smooth(method = "lm", se = T)+
  geom_jitter( alpha = .5, height = 0.04)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Performance by trial and age bin

```{r plot trials, echo = F}
ms <- d %>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  group_by(subage, trial, subid) %>%
  summarise(correct = mean(correct)) %>%
  multi_boot_standard(col = "correct")

ggplot(ms, 
       aes(x = trial, y = mean, col = subage)) +
  geom_line(aes(group= subage)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1)) + 
  geom_hline(yintercept = 0.5, lty=2)+
  theme_few() + 
  scale_colour_solarized()
```

## T-tests against chance per age bin
```{r, echo = F}
## t tests against chance
d %>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(subage) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(subage,df,t_value,p_value) %>%
  knitr::kable(digits = 3)

@@

## GLMM

```{r, include=F}
## model
library(lme4)

lm_data <- d%>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

# maximal converging model
lm <- glmer(correct ~ age_num + (1 | subid), 
              data = lm_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r, echo = F}
summary(lm)
```

# Effect of target position

Target was either the closer (inner) or further away (outer) object.

```{r, echo = F}
p5 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by( subage, targetPosition, subid) %>%
  summarise(correct = mean(correct))

p6 <- p5 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p5, aes(x = targetPosition, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p6, aes(x = targetPosition, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Target switch

In half of the trials, the target switched sides from exposure to test.

```{r, echo = F}
p7 <- d %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by( subage, targetSwitch, subid) %>%
  summarise(correct = mean(correct))

p8 <- p7 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p7, aes(x = targetSwitch, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p8, aes(x = targetSwitch, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

# Joining round 1 and round 2

```{r}
files <- dir("~/Work/MCC/git-mcc/kids_info_data/")

raw_data <- data_frame()
for (f in files) {
  jf <- paste("~/Work/MCC/git-mcc/kids_info_data/",f,sep="")
  jd <- jsonlite::fromJSON(paste(readLines(jf), collapse=""))
  date <- str_sub(jf,45,str_length(jf)-11)
  id <- as_data_frame(jd$data$data) %>% mutate(test_date = date)
  raw_data <- bind_rows(raw_data, id)
}

d_raw <-raw_data %>%
  mutate(test_date= ifelse(test_date == "18-10-10", paste(20,test_date,sep = ""),test_date),
         subid= ifelse(subid == "101018_8_inf", "181010_8_inf",subid),
         subid= ifelse(subid == "181031_3", "181031_3_inf",subid),
         test_date= ifelse(test_date == "18-XX-XX", "2018-10-10",test_date),
         test_date= ifelse(test_date == "18-09-12", "2018-09-12",test_date),
         experiment = "mcc_kids_inf_2")

log <- read_excel("../../../MCC-subject_log.xlsx", 1)%>%
  filter(Condition == "inf")%>%
  select(subid,experimenter,keep_drop,sex,dob)

d2 <- left_join(d_raw,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(dob = as.numeric(dob),
         dob = as.Date(dob,origin = "1899-12-30"),
         age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))

d_sum <- bind_rows(d%>%select(-test_date),d2%>%select(-test_date)%>%filter(subage!="2"))%>%
    filter(trial != "filler1",
         trial != "filler2")

```{r}
d_sum %>%
  group_by(subage)%>%
  summarise(n = length(unique(subid)))%>%
  knitr::kable(digits = 2)
```

```{r plot by age, echo= FALSE}

p_sum1 <- d_sum %>%
  filter(trial != "filler1",
         trial != "filler2")%>%
  group_by(control,subage, subid) %>%
  summarise(correct = mean(correct))

p_sum2 <- p_sum1 %>%
  multi_boot_standard(col = "correct")

ggplot() +
  geom_jitter(data = p_sum1, aes(x = control, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = p_sum2, aes(x = control, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  facet_wrap(~subage)+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Control",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

```{r plot by numerical age, echo = F}
p_sum3 <- d_sum %>%
  filter(trial != "filler1",
         trial != "filler2") %>%
  group_by(age_num, subid) %>%
  summarise(correct = mean(correct))

ggplot(data = p_sum3, aes(x = age_num, y = correct)) +
  geom_smooth(method = "lm", se = T)+
  geom_point( alpha = .5, width = 0.02)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  #theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized(name="Speaker Change",
                     breaks=c("false", "true"),
                     labels=c("No", "Yes"))
```

```{r, include=F}
## model
library(lme4)

lm_data_sum <- d_sum%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

# maximal converging model
lm_sum <- glmer(correct ~ age_num + (1 | subid), 
              data = lm_data_sum, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

```{r, echo = F}
summary(lm_sum)
```

>>>>

```{r import data, include = F}
###
library(tidyverse)
library(knitr)
library(langcog)
library(ggthemes)
library(jsonlite)
library(readxl)
library(brms)
library(here)
library(brmstools)
library(ggpubr)
library(rwebppl)

## importing data

files <- dir("~/Work/Spin/raw_data/kids/basic_me/")

@@

length(unique(raw_data$subid))

d_raw <-raw_data %>%
  mutate(subid = ifelse(subid == "190314_3", "190314_3_me",subid),
         subid = ifelse(subid == "190221_6_me", "190321_6_me",subid),
         subid = ifelse(subid == "192203_4_me", "190322_4_me",subid),
         subage = ifelse(subid == "190313_5_me", "3",subage),
         test_date = as.Date(test_date))

log <- read_excel("../../SPIN-subject_log.xlsx", 1)%>%
  filter(Condition == "me")%>%
  mutate(dob = as.character(dob),
         dob = ifelse(subid == "190313_7_me", "2015-11-15",dob),
         dob = ifelse(subid == "190313_11_me", "2016-04-18",dob),
         dob = ifelse(subid == "190320_8_me", "2014-04-28",dob),
         dob = as.Date(dob))%>%
  select(subid,experimenter,keep_drop,sex,dob)

d_raw%>%
  group_by(subid)%>%
  summarise(n())%>%
  filter(!(subid %in% unique(log$subid)))

d <- left_join(d_raw,log, by = "subid")%>%
  filter(keep_drop == "keep")%>%
  mutate(age_num = lubridate::time_length(difftime(test_date,dob), "years"),
         check_age = ifelse(substr(age_num,1,1) == subage, T, F))

#write_csv(d, "../data/child_basic_me.csv")

@@

```{r filler by age, echo= FALSE, warning= FALSE}
pf1 <- d %>%
      filter(trial == "train1" | trial == "train2")%>%
  group_by(subage, subid) %>%
  summarise(correct = mean(correct))

@@

ggplot() +
  geom_jitter(data = pf1, aes(x = subage, y = correct, col = subage, alpha = .2),width = .3,height = .025)+
  geom_pointrange(data = pf2, aes(x = subage, y = mean, col = subage, ymin = ci_lower, ymax = ci_upper),size = .8)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Proportion Expected Choice")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank())+
  scale_colour_solarized()
```

# Itmes by age bin

```{r plot by age, echo= FALSE}
aoa_ratings_all <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)

aoa_ratings <- aoa_ratings_all %>%
  filter(Word %in% d_raw$familiar )%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         familiar = Word)%>%
  select(familiar,mean_aoa)

p1 <- d %>%
  filter(trial != "train1",
         trial != "train2")%>%
  group_by(subage, familiar) %>%
  multi_boot_standard(col = "correct")%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(familiar = fct_reorder(factor(familiar), mean_aoa))

ggplot(data = p1, aes(x = familiar, y = mean, group=1, col = subage)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),size = .8)+
  stat_summary(aes(group = subage, col = subage),fun.y=mean, geom="line")+
  labs(x="",y="Proportion Correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  guides(alpha = F)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  scale_colour_solarized()
```

@@

```{r plot by numerical age, echo = F}
p2 <- d %>%
  filter(trial != "train1",
         trial != "train2")%>%
    left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(familiar = fct_reorder(factor(familiar), mean_aoa))

data_plot <- ggplot(data = p2, aes(x = age_num, y = correct, col = familiar, fill = familiar)) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_jitter(width = 0, height = 0.05, alpha = .6)+
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F, alpha = .5)+
  labs(x="Age",y="Proportion correct")+
  #facet_grid(~familiar)+
  ggtitle("Data")+
  guides(col = F, fill = F)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")+
  scale_fill_viridis_d(name = "familiar object")
```

```{r}
d_model <- d %>%
    filter(trial != "train1",
         trial != "train2")%>%
  mutate(age_num = scale(age_num))

model <- brm(correct ~ age_num + (1|subid) + (age_num | familiar),
                    data = d_model, family = bernoulli(),
          control = list(adapt_delta = 0.95),
          sample_prior = F,
          save_all_pars = TRUE,
          iter = 2000)

summary(model)

fixef <- as_tibble(fixef(model), rownames = "term")

ranef <-  ranef(model)

ranef_plot <- as_tibble(ranef$familiar, rownames = "familiar")%>%
  mutate(grand_intercept = fixef%>%filter(term=="Intercept")%>%pull(Estimate),
         grand_slope = fixef%>%filter(term=="age_num")%>%pull(Estimate))%>%
  group_by(familiar) %>%
  expand(Estimate.Intercept,Estimate.age_num,grand_intercept,grand_slope,age = scale(d$age_num))%>%
  mutate(y = plogis(grand_intercept + Estimate.Intercept+(Estimate.age_num+grand_slope)*age), 
         age = age+mean(d$age_num))%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(familiar = fct_reorder(factor(familiar), mean_aoa))

model_plot <- ggplot(ranef_plot, aes(x=age,y = y, col = familiar))+
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(size = 1)+
  ggtitle("Model")+
  labs(x="Age",y="Proportion correct")+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F)+ 
  scale_colour_viridis_d(name = "familiar object")

@@

```{r}
ggarrange(data_plot,model_plot, common.legend = T, legend = "right")
```

```{r}

cor_plot <- as_tibble(ranef$familiar, rownames = "familiar") %>%
  left_join(aoa_ratings)

ggplot(cor_plot, aes(x = mean_aoa, y = Estimate.Intercept))+
  geom_point()+
  geom_smooth(method = "lm")+
  xlab("Age of acquisition rating")+
  ylab("ME effect (model intercept)")+
  stat_cor(method = "pearson", label.x = 8, label.y = .99)+
  coord_fixed(ratio = 2)+
  theme_few()

@@
